% =============================================================
%  APPROVAL
% =============================================================
\chapter*{Approval}
\addcontentsline{toc}{chapter}{Approval}

\noindent I do hereby declare that the research works presented in this thesis entitled,
\textbf{``Comparing Imputation and Clustering Pipelines on Incomplete
Newspaper-Reported Suicide Data from Bangladesh: An Exploratory Study''}, are the outcome of the original
works carried out by \textbf{Mohammed Akash, Sajedul Islam, Md Arifuzzan,
Komol Chandra Devnath, and Anisul Haque Bhuiyan} under my supervision. I further
declare that no part of this thesis has been submitted elsewhere for the requirements of any
degree, award, diploma, or other purposes except for publications. I further certify that the
dissertation meets the requirements and standards for the degree of Bachelor of Science
in Computer Science and Engineering.

\vspace{0.8in}

% --- Supervisor signature ---
\noindent\hrulefill\hspace{2em}\\[2pt]
\textbf{Supervisor}\\
\textbf{Md. Mamun Hossain}\\
Assistant Professor\\
Department of Computer Science \& Engineering\\
Bangladesh University of Business and Technology\\
Dhaka, Bangladesh.

\vspace{0.6in}

% --- Chairman signature ---
\noindent\hrulefill\hspace{2em}\\[2pt]
\textbf{Chairman}\\
\textbf{Prof. Dr. Md. Ahsan Habib}\\
Professor \& Chairman\\
Department of Computer Science \& Engineering\\
Bangladesh University of Business and Technology\\
Dhaka, Bangladesh.

\clearpage
